---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# quartet

<!-- badges: start -->
<!-- badges: end -->

The quartet package is a collection of datasets aimed to help data analysis practitioners and students learn key statistical insights in a hands-on manner. It contains:

* Anscombe's Quartet
* Datasaurus Dozen
* Causal Quartet

## Installation

You can install the development version of quartet like so:

``` r
devtools::install_github("LucyMcGowan/quartet)
```

## Causal Quartet 

The goal of the `causalquartet` data set is to help drive home the point that when presented with an exposure, outcome, and some measured factors, statistics alone, whether summary statistics or data visualizations, are not sufficient to determine the appropriate causal estimate. Additional information about the data generating mechanism is needed in order to draw the correct conclusions. See [this paper](https://github.com/LucyMcGowan/writing-quartet/blob/main/manuscript.pdf) for details.


## Example

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(quartet)

ggplot(causalquartet, aes(x = x, y = y)) +
  geom_point() + 
  geom_smooth(method = "lm", formula = "y ~ x") +
  facet_wrap(~type)
```

```{r}
causalquartet %>%
  nest_by(type) %>%
  mutate(`Y ~ X` = round(coef(lm(y ~ x, data = data))[2], 2),
         `Y ~ X + Z` = round(coef(lm(y ~ x + z, data = data))[2], 2),
         `Correlation of X and Z` = round(cor(data$x, data$z), 2)) %>%
  select(-data, `Data generating mechanism` = type) %>%
  knitr::kable()
```

