---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# quartet

<!-- badges: start -->
<!-- badges: end -->

The quartet package is a collection of datasets aimed to help data analysis practitioners and students learn key statistical insights in a hands-on manner. It contains:

* Anscombe's Quartet
* Causal Quartet
* Datasaurus Dozen
* Interaction Triptych


## Installation

You can install the development version of quartet like so:

``` r
devtools::install_github("LucyMcGowan/quartet")
```

## Anscombe's Quartet

The goal of the `anscombe_quartet` data set is to help drive home the point that visualizing your data is important. Francis Anscombe generated these four datasets to demonstrate that statistical summary measures alone cannot capture the full relationship between two variables (here, `x` and `y`). Anscombe emphasized the importance of visualizing data prior to calculating summary statistics.

 * Dataset 1 has a linear relationship between `x` and `y`
 * Dataset 2 has shows a nonlinear relationship between `x` and `y`
 * Dataset 3 has a linear relationship between `x` and `y` with a single outlier
 * Dataset 4 has shows no relationship between `x` and `y` with a single outlier that serves as a high-leverage point.

In each of the datasets the following statistical summaries hold:
  * mean of `x`: 9
  * variance of `x`: 11
  * mean of `y`: 7.5
  * variance of y: 4.125
  * correlation between `x` and `y`: 0.816
  * linear regression between `x` and `y`: `y = 3 + 0.5x`
  * $R^2$ for the regression: 0.67

## Example

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(quartet)

ggplot(anscombe_quartet, aes(x = x, y = y)) +
  geom_point() + 
  geom_smooth(method = "lm", formula = "y ~ x") +
  facet_wrap(~dataset)

anscombe_quartet |>
  group_by(dataset) |>
  summarise(mean_x = mean(x),
            var_x = var(x),
            mean_y = mean(y),
            var_y = var(y),
            cor = cor(x, y)) |>
  knitr::kable(digits = 2)
```

## Causal Quartet 

The goal of the `causal_quartet` data set is to help drive home the point that when presented with an exposure, outcome, and some measured factors, statistics alone, whether summary statistics or data visualizations, are not sufficient to determine the appropriate causal estimate. Additional information about the data generating mechanism is needed in order to draw the correct conclusions. See [this paper](https://github.com/LucyMcGowan/writing-quartet/blob/main/manuscript.pdf) for details.


## Example

```{r}
#| message: false
#| warning: false

ggplot(causal_quartet, aes(x = x, y = y)) +
  geom_point() + 
  geom_smooth(method = "lm", formula = "y ~ x") +
  facet_wrap(~dataset)
```

```{r}
causal_quartet |>
  nest_by(dataset) |>
  mutate(`Y ~ X` = round(coef(lm(y ~ x, data = data))[2], 2),
         `Y ~ X + Z` = round(coef(lm(y ~ x + z, data = data))[2], 2),
         `Correlation of X and Z` = round(cor(data$x, data$z), 2)) |>
  select(-data, `Data generating mechanism` = dataset) |>
  knitr::kable()
```

## Datasaurus Dozen

Similar to Anscombe's Quartet, the Datasaurus Dozen has additional data sets where the mean, variance, and Pearson's correlation are identical, but visualizations demonstrate the large difference between datasets. This dataset is rexported from the [datasauRus](https://CRAN.R-project.org/package=datasauRus) R package.

## Example

```{r}
#| message: false
#| warning: false

ggplot(datasaurus_dozen, aes(x = x, y = y)) +
  geom_point() + 
  geom_smooth(method = "lm", formula = "y ~ x") +
  facet_wrap(~dataset)

datasaurus_dozen |>
  group_by(dataset) |>
  summarise(mean_x = mean(x),
            var_x = var(x),
            mean_y = mean(y),
            var_y = var(y),
            cor = cor(x, y)) |>
  knitr::kable(digits = 2)
```

## Interaction Triptych

This set of 3 datasets demonstrating that while the slopes estimated by a simple linear interaction model may be the same, the underlying data-generating mechanisms can be vastly different.

```{r}
ggplot(interaction_triptych, aes(x, y)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm", formula = "y ~ x", color = "cornflower blue") + 
  facet_grid(dataset ~ moderator)
```

## References

Anscombe, F. J. (1973). "Graphs in Statistical Analysis". American Statistician. 27 (1): 17–21. doi:10.1080/00031305.1973.10478966. JSTOR 2682899.

Davies R, Locke S, D'Agostino McGowan L (2022). _datasauRus: Datasets from the Datasaurus Dozen_. R package version 0.1.6, <https://CRAN.R-project.org/package=datasauRus>.

Matejka, J., & Fitzmaurice, G. (2017). Same Stats, Different Graphs: Generating Datasets with Varied Appearance and Identical Statistics through Simulated Annealing. CHI 2017 Conference proceedings: ACM SIGCHI Conference on Human Factors in Computing Systems. Retrieved from https://www.autodesk.com/research/publications/same-stats-different-graphs

Rohrer, Julia M., and Ruben C. Arslan. "Precise answers to vague questions: Issues with interactions." Advances in Methods and Practices in Psychological Science 4.2 (2021): 25152459211007368.
