---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# quartet

<!-- badges: start -->
<!-- badges: end -->

The quartet package is a collection of datasets aimed to help data analysis practitioners and students learn key statistical insights in a hands-on manner. It contains:

* Anscombe's Quartet
* Datasaurus Dozen
* Causal Quartet

## Installation

You can install the development version of quartet like so:

``` r
devtools::install_github("LucyMcGowan/quartet")
```

## Causal Quartet 

The goal of the `causalquartet` data set is to help drive home the point that when presented with an exposure, outcome, and some measured factors, statistics alone, whether summary statistics or data visualizations, are not sufficient to determine the appropriate causal estimate. Additional information about the data generating mechanism is needed in order to draw the correct conclusions. See [this paper](https://github.com/LucyMcGowan/writing-quartet/blob/main/manuscript.pdf) for details.


## Example

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(quartet)

ggplot(causal_quartet, aes(x = x, y = y)) +
  geom_point() + 
  geom_smooth(method = "lm", formula = "y ~ x") +
  facet_wrap(~type)
```

```{r}
causal_quartet %>%
  nest_by(dataset) %>%
  mutate(`Y ~ X` = round(coef(lm(y ~ x, data = data))[2], 2),
         `Y ~ X + Z` = round(coef(lm(y ~ x + z, data = data))[2], 2),
         `Correlation of X and Z` = round(cor(data$x, data$z), 2)) %>%
  select(-data, `Data generating mechanism` = type) %>%
  knitr::kable()
```

## References

Anscombe, F. J. (1973). "Graphs in Statistical Analysis". American Statistician. 27 (1): 17–21. doi:10.1080/00031305.1973.10478966. JSTOR 2682899.

Davies R, Locke S, D'Agostino McGowan L (2022). _datasauRus: Datasets from the Datasaurus Dozen_. R package version 0.1.6, <https://CRAN.R-project.org/package=datasauRus>.

Matejka, J., & Fitzmaurice, G. (2017). Same Stats, Different Graphs: Generating Datasets with Varied Appearance and Identical Statistics through Simulated Annealing. CHI 2017 Conference proceedings: ACM SIGCHI Conference on Human Factors in Computing Systems. Retrieved from https://www.autodesk.com/research/publications/same-stats-different-graphs

